### 2023-06-27 paper/paper.tex
Rewriting the method and synthetic population section

% $Id$
\documentclass[twocolumn,prd,nofootinbib]{revtex4}

\newcommand\ForInternalReference[1]{}
\newcommand\SkipForEarlyCirculation[1]{}
\newcommand\AddedResponse[1]{{\color{blue} {#1}}}
%\newcommand\SkipForEarlyCirculation[1]{#1}
\newcommand\SkipPP[1]{}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{color}
\usepackage{xspace}
\usepackage{url}
\usepackage{amsmath}
%\usepackage{adjustbox}
\usepackage{float}
\usepackage{multirow}
\usepackage{amssymb}
%
%
\usepackage{times}
%
%
%
\newcommand\optional[1]{}

%
\newcommand\E[1]{\left\langle #1\right\rangle}
\newcommand\qmstate[1]{\left|#1\right \rangle}
\newcommand\qmstateKet[1]{\left\langle#1\right|}
\newcommand\qmstateproduct[2]{\left\langle#1|#2\right\rangle}
\newcommand\qmoperatorelement[3]{\left\langle#1\left|#2\right|#3\right\rangle}
\newcommand\qmoperator[1]{{\bf #1}}
%
\newcommand\Y[1]{{{}_{#1}Y}}

\newcommand\lnL{ \ln {\cal L}}
\newcommand\lnLmarg{ \ln{\cal L}_{\rm marg}{}}
\newcommand\unit[1]{{\rm #1}}

\newcommand\rapidPEOrig{rapid\_pe1}
\newcommand\ILE{ILE}
\newcommand\editremark[1]{{\color{red} #1}}
%
%
%
\usepackage{color}
\definecolor{amber}{rgb}{1.0, 0.75, 0.0}
\definecolor{orange}{rgb}{1.0, 0.5, 0.0}
\definecolor{amaranth}{rgb}{0.9, 0.17, 0.31}
\def\fixme#1{\textcolor{red}{#1}}
\newcommand{\Richard}[1]{ {\color{blue}{#1}}}
\newcommand{\ros}[1]{ {\color{blue}{#1}}}
%

%

%
%
%
%
\graphicspath{{./figures/}}
\newcommand{\mc}{{\cal M}}
\newcommand{\Ms}{M_{\odot}}
\newcommand\LambdaTilde{\widetilde{\Lambda}}
\newcommand\DeltaLambdaTilde{\delta \widetilde{\Lambda}}
%
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\simlt{\lower.5ex\hbox{\ltsima}}
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\simgt{\lower.5ex\hbox{\gtsima}}

\newcommand\prx{Phys.~Rev.~X}
\def\aj{Astronomical Journal}                 %
\def\apj{Astrophysical Journal}                %
\def\apjl{Astrophysical Journal}             %
\def\pasj{PASJ}
\def\apjs{ApJS}              %
\def\mnras{MNRAS}            %
\def\prd{Phys.~Rev.~D}       %
\def\prl{Phys.~Rev.~Lett}    %
\def\cqg{Class.~Quant.~Grav.~}%
\def\araa{ARA\&A}             %
\def\nat{Nature}              %
\def\aap{A\&A}                %
\def\aapr{A\&A~Rev.~}    %
\def\pasp{PASP}    %
\def\sovast{Soviet Ast.}
%
%

\newcommand{\IMRPD}{\textsc{IMRPhenomD}\xspace}
\newcommand{\IMRPDT}{\textsc{IMRPhenomD\_NRTidal}\xspace}
\newcommand{\IMRP}{\textsc{IMRPhenomPv2}\xspace}
\newcommand{\SEOBP}{\textsc{SEOBNRv3}\xspace}
\newcommand{\SEOBA}{\textsc{SEOBNRv4}\xspace}
\newcommand{\SEOBAROM}{\textsc{SEOBNRv4\_ROM}\xspace}
\newcommand{\NRSur}{NRSur7dq2\xspace}
\newcommand{\TEOB}{SEOBNRv4T\xspace}
\newcommand{\Resum}{TEOBResumS\xspace}
\newcommand\RIFT{RIFT}
\newcommand{\Taylor}{TaylorF2\xspace}
\newcommand\PaperDetection{\underline{LVC-detect}\cite{DiscoveryPaper}}
\newcommand\PaperPE{\underline{LVC-PE}\cite{PEPaper}}
\newcommand\PaperTestGR{\underline{LVC-TestGR}\cite{TestingGRPaper}}
\newcommand\PaperPENRMethods{\underline{PE+NR-Methods}\cite{gwastro-PENR-Methods-Lange}}
\newcommand\PaperAstro{\underline{LVC-Astro}\cite{AstroPaper}}
\newcommand\PaperBurst{\underline{LVC-Burst}\cite{BurstPaper}}
\newcommand\PaperRates{\underline{LVC-Rates}\cite{RatesPaper}}
\newcommand\PaperStochastic{\underline{LVC-Stochastic}}
\newcommand\PaperSEOBNRvthree{\underline{LVC-SEOBNRv3}\cite{SEOBv3Paper}}

\def\RIT{Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, New York 14623, USA}

\begin{document}
\renewcommand{\arraystretch}{1.5}
\title{Population Inference of Non-Spinning Eccentric Binary Black Holes}
%\author{R. O'Shaughnessy}
\affiliation{\RIT}
%\author{M. Zeeshan}

\affiliation{\RIT}
\begin{abstract}
Yet to work. 
\end{abstract}
\maketitle

\section{Introduction}
Big Picture
\begin{itemize}
    \item GWs are important
    \item eccentricity is a signature that can be measured with GWs
    \item it is a unique signature of strong events
\end{itemize}
Why Population, smaller picture
\begin{itemize}
    \item population inference of eccentricity is important.
    \item Some events may have unabmoguis eccentricity (refs to PE, Richard will send them), but many will have only faint signatures or none at all.
    \item critical for astrophysics to acces distribution of eccentricities and any subpopulation
\end{itemize}
Previous work
\begin{itemize}
    \item What other people have done with eccentricity, 
    \item cite any previous work related to eccentricity with population.
    \item mostly it's real events or high mass events
    \item what sort of eccentricity we expect to appear
\end{itemize}
\begin{itemize}
    \item Explain, what and how you are doing here.
    \item assesses how well the eccentricity of the population  can be inferred using low mass BHS where eccentricity measurement should be much more precise.
    \item for simplicity using a proof of concept population model to emphasize eccentric inference
\end{itemize}

We discussed the various formation scenario of EBBHs in the previous section; each one has a distinct signature in the GW signals. Hence, it's trivial to observe the effect of each scenario or the mixed effect on the merging population. It leads us to make the population model, which considers each scenario and allows distinct merger rates for them.

Most of the eccentric binary black holes (EBBHs) are driven by low masses $(10M_\odot-30M_\odot)$ and if any component of the binary has mass less than $30 M_\odot$ then only $10\%$ of them are able to maintain eccentricity near the last stable orbit.  \cite{mass_ecc_limit_2018}. Therefore,  in this study, we  will consider binaries with non-spinning $(\chi_i = 0)$, eccentric $(0-0.5)$ and moderate mass range $(10M_\odot-50M_\odot)$. In addition, we will use mass ratio $q=m_1/m_2$ with condition $m_1>m_2$ and total mass $M=m_1+m_2=100M_\odot$.

\section{Methods}
\label{sec:methods}

The coalescing BBH can be completely described by three intrinsic and seven extrinsic parameters. The intrinsic parameters such as mass of binary component ($m_i$), spin $( \chi_i)$, and eccentricity $\epsilon$  are subject to the orbital evolution of the binary. The extrinsic parameters are orientation (orbital phase, polarization, and inclination), sky location (right ascension and declination), luminosity distance, and coalescence time depend on the observer.

\subsection{Hierarchical Bayesian Modeling (HBM)}

We use HBM to constrain the population models with gravitational wave signals. It is also a good choice to consider the sensitivity of LIGO-VIRGO-KAGRA (LVK) and uncertainty in each detection.

In HBM, we have the N number of discrete detections. Those detections provide merger data denoted as $d_1,d_2,d_3,...,d_N$ where each $d_i$ shows a BBH merger. Each $d_i$ has mass, spin, and eccentricity properties. These properties, often called parameters, are denoted by $\lambda_1,\lambda_2,\lambda_3,...,\lambda_i$. Each parameter has its uncertainty, and we express it by the probability of the data given the parameter value. We also refer to it as the likelihood function $\mathcal{L}(\lambda)=p(d|\lambda)$ of one event. A given value of $\lambda$ is obtained from a mathematical model commonly known as a waveform to calculate a likelihood. Once you have a likelihood function, you may use a uniform prior or any other prior to find a posterior probability using the Bayes theorem as given in Eq. \ref{eq:Bayes_ind}

\begin{equation}
\label{eq:Bayes_ind}    
p(\lambda|d) \propto p(d|\lambda) p(\lambda).
\end{equation}

This posterior probability will constrain the properties of each binary, such as mass, spin, and eccentricity. We may infer those parameters using rapid parameter inference on gravitational wave sources via iterative fitting (RIFT): an open source code for parameter estimation (PE) of the binary sources \cite{rift_2018}.

%\begin{itemize}
%    \item We have N discrete detection.
%    \item each detection has data $d_1,d_2,d_3,...,d_N$
%    \item each data point $d_N$ has properties denoted by $\lambda_1,\lambda_2,...,\lambda_i$
%    \item  $ \lambda_i$ are parameters such as mass, spin, eccentricity, and location.
%    \item Each parameter has its uncertainty.
%    \item This uncertainty is described by the probability of the data given the parameter value. We call it the likelihood function of one single event. $\mathcal{L}(\lambda) = p(d|\lambda)$
%    \item the waveform gives the parameter value.
%    \item waveform is a mathematical function.
%    \item Once you have a likelihood function, you can use prior( usually it's uniform: which gives equal probability to each event) to find the posterior. $p(\lambda|d) \propto p(d|\lambda) p(\lambda)$

%\end{itemize}

\subsection{Population Inference}

Now having the PE of individual sources, we will follow the Bayesian framework for population inference. The likelihood of a population parameter $\Lambda$, also considered as uncertainty in $\Lambda$ is equivalent to the probability of the individual sources given the population parameter $\Lambda$ is written as follows.

\begin{equation}
\label{eq:likelihood_pop}    
\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)
\end{equation}

We will use the likelihood given in Eq. \ref{eq:likelihood_pop} in the Bayes theorem given in Eq. \ref{eq:Bayes} to find the posterior probability.

\begin{equation}
\label{eq:Bayes}    
p(\Lambda|d_1,d_2,...,d_N)= \frac{p(\Lambda)p(d_1,d_2,...,d_N|\Lambda)}{p(d_1,d_2,...,d_N)},
\end{equation}

where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization constant or also known as evidence.

To conduct our analysis for mass and eccentricity distribution, we will use the inhomogeneous Poisson process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$.

\begin{equation}
\label{eq: likelihood}
\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda \ell_n(\lambda) \mathcal{R} p(\lambda|\Lambda),
\end{equation}

where $\mu(\mathcal{R},\Lambda)$ is the expected number of detection under the given population parametrization $\Lambda$ with the overall rate $\mathcal{R}$. $\ell_n(\lambda)=p(d_n|\lambda)$ is the likelihood of the data $d_n$ given binary parameter.
Finally, we will get our posterior as $p(\mathcal{R},\Lambda | D)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$ by choosing a prior $p(\mathcal{R},\Lambda)$.

These calculations are analytically intractable and must be performed numerically. Specifically, we will use Goodman and Weare's affine invariant Markov chain Monte Carlo (MCMC) \cite{mcmc_paper} to find the posterior distribution of population parameters. This method draws samples from the targeted distribution for $\Lambda$, in our case, it's the power law model given in Eq. \ref{eq:plawg}, then compares it with the given data (collection of individual events) and stores the best-fit sample. We may iterate this as we need and store multiple sample values untill they converge. 
This method is in a Python package called EMCEE \cite{emcee_paper}.

\subsection{Volume Time (VT) Estimation}

To make our study realistic, including the sensitivity of the LVK instruments is significant. This sensitivity is defined by VT estimation. $V$ is the characteristic volume with units $Gpc^{-3}yr^{-1}$, which refers to the possible detection region in the sky for the LVK \cite{Volume_1993}, and $T$ is the time duration of making observations at this sensitivity.
Existing LVK instruments' sensitivity depends only on the mass; therefore, we will get the same VT calculated previously \cite{Dan_2019} for non-spinning, non-eccentric, and no-precessing binaries. Hence, we briefly explain the calculations, see \cite{Dan_2019} for details.

The Eq. \ref{eq:volume} calculates the orientation averaged sensitive volume \cite{Abbott_2016,richard2010volume}

\begin{equation}
\label{eq:volume}
V(\lambda) = \int P((<D(z))/D_n(\lambda))\frac{dV_c}{dz}\frac{dz}{1+z}
\end{equation}    
where $D(z)$ is the luminosity distance for redshift $z$, $V_c$ is the comoving volume. Finally, to compute the average number of detection, we will use Eq. \ref{eq:mu} and keep in mind that the merger will be accepted only if the signal-to-noise ratio exceeds 8 \cite{SNR_2010}.

\begin{equation}
\label{eq:mu}
  \mu(\mathcal{R},\Lambda) = \int(VT)\lambda \mathcal{R}p(\lambda|\Lambda)d\lambda ,
\end{equation}
where $p(\lambda|\Lambda)$ is the probability density function for a random binary in the universe to have intrinsic parameter $\lambda$. Keep in mind that $\lambda$ is equal to all intrinsic and extrinsic parameters.

\subsection{Power law Model}
There are various weak and pure phenomenological population models proposed in previous studies \cite{2016PRXAbbot_BBH_model,2017FishBach_BBH_model,2018talbot_bbh_model}. However, our analysis used the pure truncated power law defined in \cite{2016PRXAbbot_BBH_model,2017FishBach_BBH_model} and Gaussian eccentricity. This model computes the intrinsic probability of $m_1$, $m_2$, and $\epsilon$.  
We also assume that non-zero probability density only exists for $m_{min}\leq m_2 \leq m_1 \leq m_{max}$ and for total mass $M_{max}=m_1+m_2 = 200 M_\odot$. The condition of the total mass is only because of the limitations of the detectors towards the higher masses \cite{2016PRXAbbot_BBH_model}. The generalized form of the truncated power-law model with parameters $\Lambda \equiv  (\alpha, \mathcal{R}, k_m, m_{min}, m_{max}, \sigma_\epsilon, M_{max})$ and random variable $m_1$, $m_2$, and $\epsilon$ has the functional form in Eq. \ref{eq:plawg} within provided mass limit.

\begin{align}
\label{eq:plawg}
p(m_1,m_2,\epsilon) = &C(\alpha,k_m,m_{min},m_{max},M_{max},\epsilon)  
\nonumber \\ & \sqrt{\frac{2}{\pi}} \frac{(m_2/m_1)^{k_m} m_1^{-\alpha} e^{-(\epsilon/\sqrt{2}\sigma_\epsilon)^2}}{(m_1-m_{min})\sigma_\epsilon},
\end{align}

where $\alpha$ is the power law index, $\mathcal{R}$ is the merger rate, $m_{min}, m_{max}$ are the minimum and maximum masses of the binary components in the population, and $\sigma_e$ is the orbital eccentricity distribution. The Eq. \ref{eq:plawg} represent a truncated power law for primary mass $m_1$ with index $-\alpha$ and conditional power law distribution $p(m_2|m_1)$ for secondary mass $m_2$ using simple power law, and Gaussian distribution for orbital eccentricity $e$. 
For our analysis, we defined a constant of integration equal to $\int_V dm_1 dm_2 d\epsilon p(m_1,m_2,\epsilon) = 1$.  Our detectors are sensitive to high-mass BBHs, particularly $M_{max}\geq 200 M_\odot$, therefore, we will use $k_m=0$ throughout the studies. As a result, we have our reduced form of the truncated power law in Eq. \ref{eq:plaw}

\begin{align}
\label{eq:plaw}
p(m_1,m_2,\epsilon) = \sqrt{\frac{2}{\pi}} \frac{ m_1^{-\alpha}  e^{-(\epsilon/\sqrt{2}\sigma_\epsilon)^2}}{(m_1-m_{min})\sigma_\epsilon}
\end{align}

\section{Synthetic Population}

We have created four synthetic populations by choosing the power law parameters $\alpha = -1$, $m_{min} = 10$, $m_{max}=50$. Each population has the same values for $\alpha$, $m_{min}$, and $m_{max}$ except $\sigma_\epsilon$. We selected four values for $\sigma_\epsilon = 0.05, 0.1, 0.15, 0.2$, and  generated synthetic population containing $N=10000$ sources.

\subsection{Synthetic population with eccentricity}

After generating four populations, we find the probability for each event to be detected by computing the VT of each source. Finally, we did the weighting based on these VTs and randomly picked $N=100$ sources from each population to perform our analysis. Our weighted populations with $\sigma_\epsilon=0.05$ and $\sigma_\epsilon=0.2$ are given in Fig. \ref{fig:pop3d0.05_0.2}.  The complete description of each EBBHs population is given in Table \ref{tab:pop_prop}.

\begin{figure*}[]
\includegraphics[width=0.45\textwidth]{paper/figures/pop3d_0.05.png}
\includegraphics[width=0.45\textwidth]{paper/figures/pop3d_0.2.png}
\caption{\label{fig:pop3d0.05_0.2} Synthetic Population of EBBHs. The left figure represents the population for $\sigma_\epsilon=0.05$ and right figure represents the population for $\sigma_\epsilon=0.2$}
\end{figure*}

\begin{table*}[]
    \centering
    \begin{tabular}{c|cccccc}
        \hline \hline
        Populations & $m_{1_{min}} [M_\odot] $ & $m_{1_{max}} [M_\odot]$ & $m_{2_{min}} [M_\odot]$ & $m_{2_{max}} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
        First  & 12.19 & 49.89 & 10.17 & 46.67 & 0.0004 & 0.0494\\ \hline
        Second  & 15.42 & 49.79 & 10.46 & 47.59 & 0.0003 & 0.0999\\ \hline
        Third  & 16.17 & 49.83 & 10.69 & 47.52 & 0.0002 & 0.148\\ \hline
        Forth  & 12.52 & 49.89 & 10.08 & 49.51 & 0.0012 & 0.1984\\ \hline
    \end{tabular}
    \caption{Population Properties of EBBHs}
    \label{tab:pop_prop}
\end{table*}
To make our study more realistic, we must add the measurement error in each source. The chirp mass and symmetric mass ratio are well-constrained compared to the primary and secondary masses of BBHs. Therefore, we compute the chirp mass and symmetric mass ratio of each event in the population by using the following relations.

\begin{equation}
    M_c^T = \frac{(m_1 m_2)^{3/5}}{(m_1+m_2)^{1/5}},
\end{equation}

\begin{equation}
    \eta^T = \frac{(m_1 m_2)}{(m_1+m_2)^2},
\end{equation}
where $M_c^T$ and $\eta^T$ are found using the primary and secondary masses of each event generated by the power law model.
Furthermore, using the following relations, We add the measurement errors in the $M_c^T$ and $\eta^T$.
\begin{equation}
    M_c = M_c^T\left( 1+\alpha (r_0+r ) \frac{12}{\rho}\right),
\end{equation}

\begin{equation}
\eta = \eta^T\left( 1+0.03 (r_0'+r') \frac{12}{\rho}\right),   
\end{equation}
where $r_0$ and $r_0'$ are the random numbers drawn from the standard normal distribution, which will shift the mean of the $M_c$ and $\eta$ distribution with respect to $M_c^T$ and $\eta^T$. The $r$ and $r'$ are the independent and identically distributed arrays of those randomly generated numbers to spread the distribution. The measurement uncertainty is inversely proportional to signal-to-noise ratio $\rho$, drawn from the distribution $p(\rho) \propto \rho^{-4}$, which holds for isotropically distributed sources in a static universe, subject to the threshold $\rho\geq 8$ for detection. The $\alpha =0.1$ is used for the scaling inspired by analyses of mock data with the LALINFERENCE pipeline \cite{alpha_error_2015} and includes the impact of correlation with parameters describing arbitrary remnant spins.

Finally, after adding the measurement errors in the $M_c$ and $\eta$, we will convert them back to $m_1$ and $m_2$ to perform our analysis. We used the following relation for conversion, and it will provide the masses based on the condition $m_1\geq m_2$.

\begin{equation}
    m_1 = \frac{1}{2} M_c \eta^{-3/5} (1+\sqrt{\eta_v}),
\end{equation}

\begin{equation}
    m_2 = \frac{1}{2} M_c \eta^{-3/5} (1-\sqrt{\eta_v}), 
\end{equation}
where $\eta_v = 1-4\eta$, we kept the samples with non-negative values and ignored the negative samples to avoid the square root issues. We also added the absolute error in the eccentricity by using the truncated normal distribution scaling at $0.05$ to keep the eccentricity positive.

\subsection{Synthetic population without eccentricity}

To compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we created four more populations by scaling the existing populations using the Eq. \ref{eq:scaling} defined in \cite{2021_scaling_paper}. The Eq. \ref{eq:scaling} removes the eccentric component from the population, scales the masses, and eventually, we get non-eccentric populations to compare with the existing eccentric population.
\begin{align}
\label{eq:scaling}
M^{ecc} = \frac{M}{(1-\frac{157}{24}\epsilon^2)^{3/5}}
\end{align}

\begin{table*}[]
    \centering
    \begin{tabular}{c|cccc}
        \hline \hline
        Populations & $m_{1_{min}} [M_\odot] $ & $m_{1_{max}} [M_\odot]$ & $m_{2_{min}} [M_\odot]$ & $m_{_2{max}} [M_\odot]$ \\ \hline
        First & 12.09 & 49.6 & 10.17 & 46.27 \\ \hline
        Second & 15.26 & 49.77 & 10.3 & 47.56 \\ \hline
        Third & 15.96 & 49.5 & 10.62 & 46.60  \\ \hline
        Forth & 11.86 & 49.76 & 9.67 & 46.99  \\ \hline
    \end{tabular}
    \caption{Population Properties of NEBBHs}
    \label{tab:popscl_prop}
\end{table*}

\begin{figure*}

\includegraphics[width=0.45\textwidth]{paper/figures/pop2d_0.05.png}
\includegraphics[width=0.45\textwidth]{paper/figures/pop2d_0.2.png}
\caption{\label{fig:pop2d_0.05_0.2} The left figure shows the primary mass vs secondary mass of the EBBHs and NEBBHs with $\sigma_\epsilon =0.05$. The right figure shows the same plot but for $\sigma_\epsilon=0.2$}

\end{figure*}

To get the non-eccentric population, we apply the Eq. \ref{eq:scaling} on the population given in the table. \ref{tab:pop_prop}. After applying the scaling equation, we get the new populations with scaled masses, giving us the new minimum and maximum values for the primary and secondary masses given in table \ref{tab:popscl_prop}.

The significant effect after scaling is the mass shift, which can be observed in Fig. \ref{fig:pop2d_0.05_0.2}. In this figure, the left-hand side shows the mass shift of the first population generated with smaller $\sigma_\epsilon =0.05$, which leads to a lesser mass shift. However, on the right hand of Fig. \ref{fig:pop2d_0.05_0.2}, one can see the significant mass shift after removing the eccentric component.  This also reflects that we may miss the various sources by ignoring the effect of eccentricity and we notice that sources start missing, which has eccentricity $\epsilon>0.38$.

\section{Population Inference}
%\renewcommand{\arraystretch}{2}
\begin{table*}[]
    \centering
    \begin{tabular}{c|ccccc}
        \hline \hline
       Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ & $\sigma_\epsilon$ \\ \hline
      First& $1.42^{+0.58}_{-2.64}$ & $-0.88^{+3.17}_{-1.89}$ & $8.36^{+7.49}_{-3.51}$ & $58.44^{+10.21}_{-24.31}$ & $0.30^{+0.54}_{-0.25}$ \\ \hline
      Second & $1.36^{+0.63}_{-2.65}$ & $-0.93^{+3.29}_{-1.92}$ & $8.71^{+7.36}_{-3.81}$ & $57.84^{+10.81}_{-25.53}$ & $0.33^{+0.51}_{-0.25}$ \\ \hline
      Third & $1.39^{+0.58}_{-2.69}$ & $-0.69^{+2.98}_{-1.95}$ & $9.21^{+6.31}_{-3.99}$ & $57.85^{+10.73}_{-27.04}$ & $0.33^{+0.51}_{-0.23}$  \\ \hline
      Forth & $1.27^{+0.72}_{-2.68}$ & $-0.90^{+3.30}_{-1.98}$ & $8.63^{+7.85}_{-3.72}$ & $57.76^{+11.03}_{-25.82}$ & $0.37^{+0.47}_{-0.23}$  \\ \hline
    \end{tabular}
    \caption{Population Inference for EBBHs}
    \label{tab:inference_EBBHS}
\end{table*}

\begin{table*}[]
    \centering
    \begin{tabular}{c|cccc}
        \hline \hline
        Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ \\ \hline
      First & $1.85^{+0.14}_{-1.23}$ & $-1.37^{+1.43}_{-0.94}$ & $7.54^{+3.68}_{-3.04}$ & $63.29^{+8.00}_{-13.08}$  \\ \hline
      Second & $1.83^{+0.14}_{-1.38}$ & $-1.12^{+1.47}_{-0.91}$ & $7.69^{+3.88}_{-3.15}$ & $61.91^{+8.59}_{-13.87}$  \\ \hline
      Third & $1.84^{+0.15}_{-1.55}$ & $-0.38^{+1.03}_{-0.96}$ & $8.07^{+3.44}_{-3.27}$ & $61.70^{+8.49}_{-24.26}$   \\ \hline
      Forth & $1.86^{+0.16}_{-1.85}$ & $-0.31^{+1.17}_{-1.14}$ & $6.99^{+8.77}_{-2.58}$ & $61.15^{+8.72}_{-16.84}$  \\ \hline
    \end{tabular}
    \caption{Population Inference for NEBBHs}
    \label{tab:inference_NEBBHS}
\end{table*}
we can discuss the corner plots here.

\begin{figure*}

\includegraphics[width=0.95\textwidth]{paper/figures/cor_0.05.png}
\caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHS for $\sigma_\epsilon=0.05$}}

\end{figure*}
Same population.
\begin{figure*}

\includegraphics[width=0.95\textwidth]{paper/figures/cor_0.2.png}
\caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHs for $\sigma_\epsilon=0.2$}}

\end{figure*}

Need to write it later.

% \begin{align} \label{eq:strain_mode}
% h(t,\vartheta,\phi;\bm{\lambda}) = 
% \sum_{\ell=2}^{\infty} \sum_{m=-\ell}^{\ell} \frac{D_{\rm ref}}{D} h^{\ell m}(t;\bm{\lambda}) \Y{-2}_{\ell m} \left(\vartheta, \phi \right) \, ,
% \end{align}

% \begin{widetext}
% \begin{align}
% \ln {\cal L}(\bm{\lambda}, \theta) 
% &= (D_{\rm ref}/D) \text{Re} \sum_k \sum_{\ell m}(F_k \Y{-2}_{\ell m})^* Q_{k,lm}(\bm{\lambda},t_k)\nonumber \\
% &   -\frac{(D_{\rm ref}/D)^2}{4}\sum_k \sum_{\ell m \ell' m'}
% \left[
% {
% |F_k|^2 [\Y{-2}_{\ell m}]^*\Y{-2}_{\ell'm'} U_{k,\ell m,\ell' m'}(\bm{\lambda})
% }
% % \right. \nonumber \\ & \left.
%  {
% +  \text{Re} \left( F_k^2 \Y{-2}_{\ell m} \Y{-2}_{\ell'm'} V_{k,\ell m,\ell'm'} \right)
% }
% \right]
% \label{eq:def:lnL:Decomposed}
% \end{align}
% \end{widetext}

% \begin{eqnarray}
% {\cal L}_{\rm margT} \equiv  \int {\cal L} \frac{dt}{T}
% \label{eq:lnL:tmarg}
% \end{eqnarray}

% \begin{table*}
% \begin{tabular}{lrr|ccccc|rr}
% Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
% $\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
%   &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
% % ~/parse_report.sh profile_nogpu_pcdev13.log | more
% CPU & 16384 & $\pm 2 $ & 20 & 2.4 &&540 & 20 &  690  \\ 
%        & 4096 & $\pm 2 $ &   20  &&&& 20 \\ \hline
% % ./parse_report.sh 20190130-profile_nogpu_HM_pcdev13.log  | more
% % /parse_report.sh ./profile_nogpu_HM_lowres_pcdev13.log 
% %    setup time: PrecomputeLikelihoodTerms, includes waveform generation. 
% %   evaluation: FactoredLogLikelihodTimeMarginalized Divide by actual number of calls, since not a block!
% %    
% CPU & 16384 & $\pm 2,\pm 1 $ & 20 & 1.5 && 680 & 20 &  1060  \\ 
%        & 4096 & $ \pm 2, \pm 1 $ &   20 &&&& 20  \\ \hline

%GPU (a) & 16384 & $\pm 2 $  & 20  & & && & 270 \\
%            & 4096 &$\pm 2 $  &  20 &  & & & & 45 \\ \hline

%GPU (b) & 16384 & $\pm 2$ & 20  & 1.8 & 1 & 0.85& 20 &28 & 15\\
%        & 4096 & $\pm 2$  & 20 & $1.2 $ &  1  & 0.75 & 20  & 25\\ \hline

% GPU (b) & 16384 & $\pm 2, \pm 1$ & 20 & 1 && 4.2 & 20  & 38  \\
%        & 4096 & $\pm 2, \pm 1$ & 20 & 1&& 2.5  & 20 & 35 & \\ \hline

% GPU (c) & 16384 & $\pm 2 $  & 20  &6  & & 18&  58&160 &  \\
%             & 4096 &$\pm 2 $  &  20 & 3.7 & & 11  & 58  & 140 & $\simeq 50$ \\
% \end{tabular}
% Compute node at LIGO-WA
% \caption{\label{tab:CostBreakdown}\textbf{Profiling performance: Binary black holes}: Evaluation costs for the
%   marginalized likelihood on default
%   hardware, for a two-mode system $(l,m)=\pm 2$ analyzing $T=8\unit{s}$ of data with a massive binary black hole
%   $m_1=35 M_\odot,M_2=30 M_\odot$.  The last column indicates peak GPU utilization.
% }
%\end{table*}

% \subsection{Binary black hole analysis}
% \label{sec:sub:BBHFull}

% \begin{figure*}

% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_m2.png}
% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_cum.png}
% % python plot_mean_variance.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency_meanVar.dat  --output bbh_zerospin_lnL_meanVar
% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_meanVar.png}
% % python plot_convergence.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency.dat
% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_converge.png}
% \caption{\label{fig:BBH:MultiIterate}\textbf{Convergence of BBH analysis: Zero spin}: Results for marginal posterior distributions
%   of our fiducial synthetic binary black hole.  Solid contours show credible intervals; solid one-dimensional distributions
%   show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
%   value of the underlying marginalized likelihood evaluations.   
% %The corresponding dotted curves show an analysis using   only the $m=\pm 2$ modes \editremark{perform both} 
% \emph{Left panel } Posterior distribution
%   over  $\mc$ and
%   $\delta=(m_1-m_2)/M$.    \emph{Right panel}: Marginal 1d CDFs of $\mc$, showing convergence.
% \emph{Bottom left}: Mean and variance of  \AddedResponse{the array $\ln{\cal L}_{\rm marg}(\bm{\lambda}_j)$  for
% $j=1,2,\ldots N_{\rm eval}$ indexing all candidate sets of intrinsic parameters $\bm{\lambda}_j$ performed in that iteration},  showing that after the
% first iteration the
% candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
% \emph{Bottom right panel}: The estimated evidence $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
% error budget, Monte Carlo error is not shown.
% }
% \end{figure*}

\section{Conclusions}
\label{sec:conclude}
In Progress,

\begin{acknowledgements}
We thank our anonymous referee for the helpful feedback.

\end{acknowledgements}

\appendix
In Progress,

%\bibstyle{unsrt}
\bibliography{paperexport,LIGO-publications}
\end{document}